---
title: "Homework.2.Chart"
format: html
editor: source
---

## Loading packages

```{r}

library(ggplot2)
library(dplyr)
install.packages("hrbrthemes")
library(hrbrthemes)

```

## Creating a dataframe


```{r}
data <- data.frame(name=c("north","south","south-east","north-west","south-west","north-east","west","east"), val=sample(seq(1,10), 8))

data
```

#### Making a scatterplot of color depending on species

```{r}

ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, color=Species)) + 
    geom_point(size=6) +
    theme_ipsum()

```

#### Inseting a new dataset from excel


```{r}

library(readxl)
Homework_2 <- read_excel("~/Desktop/Homework.2.xlsx")
View(Homework_2)
```

#### Plotting Homework.2 Data

```{r}

ggplot(Homework_2, aes(x=Month, y=Concentration, color=NutrientType)) +
         geom_point(size=6) + theme_ipsum()

```

#### Notes on this plot

* First, I tell R what dataset I'll be using (homework2)
* Then, I define the x values, the y values and the category that I want to be resembled by different colors
* Then I define the size and the color theme with the geom_point function